---
title: "panama_cleaning"
author: "Jake Eisaguirre"
date: "7/6/2022"
output: html_document
---

# packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

# librarian downloads, if not already downloaded, and reads in needed packages

librarian::shelf(tidyverse, here, janitor, lubridate, parsedate)

```

# read in the data
```{r}

pan_data <- read_csv(here("data", "panama_data.csv")) %>% 
  clean_names() # <- this function attempts to clean column names

```

# I am now going to split the data apart. The first data frame will contain meta_data and environmental variables
```{r}

md_enviro <- pan_data %>% 
  select(c(1:35))

```


### Now I will be making the `nofrog` column binary (1 = present and 0 = absent) and calling it `frog_presence`
```{r}

enviro <- md_enviro %>% 
  mutate(frog_presence = ifelse(!nofrog == "No frogs", 1, 0 )) %>% 
  mutate_at(vars(frog_presence), ~replace_na(., 1))




# now remove old column and re-position frog_presence column

binary_environ <- enviro %>% 
  select(!c(nofrog)) %>% 
  relocate(frog_presence, .before = observers)
  

```

### I am now going to format the `date` for ISO 8601
```{r}

date_environ <- binary_environ %>% 
  mutate(date = parse_date(date))


```

### I am going to rename columns to include units
```{r}

var_environ <- date_environ %>% 
  rename(duration_minutes = surv_length,
         elevation_m = elevation, # double check this is correct
         wind_speed_kph = wind_speed, # need to double check kph, I see some showing "0.4/s"
         air_temp_c = air_temp, # double check this is correct
         water_temp_c = water_temp, # double check this is correct
         dis_o2_insertunits = dis_o2, # insert units
         conductivity_insertunits = conductivity, #insert units
         tds_insertunits = tds, # insert units
         salinity_insertunits = salinity, # insert units
         cloud_cover_percent = cloud_cover,# double check this is correct
         precip_mm = precip) # double check this is correct

```

### make all text lower case and select important survey md
```{r}

survey_meta_data <- mutate_all(var_environ, str_to_lower) %>% 
  select(c((1:3), (6:15), (21:35)))

```


### save survey meta data csv
```{r}

write.csv(survey_meta_data, here("data", "survey_meta_data.csv"))

```


### Pull out site data for sites table and keep a survey table
```{r}

site_data <- survey_meta_data %>% 
  select(c(sort_id, merge_id, survey_id, region, site, utmz, utme, utmn, var_m, elevation_m))

```


### save sites data csv
```{r}

write.csv(site_data, here("data", "site_data.csv"))

```


# Next we will clean up the biological data
```{r}

surv <- pan_data %>% 
  select(c((1:3),(36:72)))

```


### create binary `infection_status` (infected = 1, unifected = 0) column and combine the two binary columns `infected` and `unifected`
```{r}

infec_status_bio <- surv %>% 
  mutate(infection_status = ifelse(infected == 1, 1, 0))

# now remove old columns and re-position infection_status column

infec_status_bio <- infec_status_bio %>% 
  select(!c(infected, uninfected)) %>% 
  relocate(infection_status, .before = pcr)

```

### rename columns to include units
```{r}

col_clean_bio <- infec_status_bio %>% 
  rename(count = quantity,
         capture_tx_loc = capture_loc, # double check this is rigth interpetation 
         body_temp_c = body_temp, # double check this is the right unit
         subs_temp_c = subs_temp, # double check this is the right unit
         svl_units = svl, # need units
         frog_mass_g = frog_mass, # double check this is the right unit
         life_stage_moredetail = life_stage0, # need more detail
         microhab_moredetail = microhab0, # need more detail
         location_moredetail = location0, # need more detail
         salinity_moredetail = salinity0,
         photo_captured = photo) # need more detail

```

### Now we want to pull table `col_clean_bio` apart for the 2 different detection types: `capture` and `call`/`visual`
```{r}

vis_call_bio <- col_clean_bio %>% 
  filter(detection_type %in% c("call", "visual")) %>% 
  select(!c(frog_mass_g, dead, infection_status, pcr, bdswab, gen_samp, bacteria_swab, am_ps, photo_check_needed, notes_species_merge, 
            notes_pc_rdata, missing_dat, swab_id))

capture_bio <- col_clean_bio %>% 
  filter(detection_type == "capture")



```


